---
title: "ShortStamen_LSMs"
author: "Sophia Buysse"
date: "1/26/2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts = list(width.cutoff=75), tidy = TRUE)
```

# Set up Model \

```{r, warning = FALSE}
## load packages 
require(emmeans)
require(lme4)
require(dplyr)
#require(lsmeans)

## load data
Full <- read.csv("C:/Users/Sophie/Michigan State University/Conner, Jeffrey - SophieAnalyses/SpanishMasterDataset_ByFlwr.csv")
Sequenced <- read.csv("C:/Users/Sophie/Michigan State University/Conner, Jeffrey - SophieAnalyses/SpanishMasterDataset_Sequenced.csv")
PopMetaData <- read.csv("C:/Users/Sophie/Michigan State University/Conner, Jeffrey - SophieAnalyses/SpanishMasterDataset_PopMetaData.csv", header=TRUE)

## format columns as factors for both datasets

Full$Population <- as.factor(substr(x=Full$Population, 1, 3))
Full$Matriline <- as.factor(Full$Matriline)
Full$Plant <- as.factor(toupper(Full$Plant))
Full$PlantingBlock <- as.factor(Full$PlantingBlock)
Full$Tube <- as.factor(Full$Tube)
Full$Notes <- NULL
str(Full)

Sequenced <- Sequenced[,1:9]
Sequenced$Population <- as.factor(Sequenced$Population)
Sequenced$Line <- as.factor(Sequenced$Line)
Sequenced$Rep <- as.factor(toupper(Sequenced$Rep))
# Plant is equivalent to Rep 
Sequenced$Tube <- as.factor(Sequenced$Tube)
str(Sequenced)
```

Start with the means for all plants that were phenotyped (Full dataframe)

```{r}

## create a fully nested random model
colnames(Full)
# the important ones are Pop, Line, Rep, Tube to mimic what Jeff used
# Population
# Matriline
# Plant
# Tube = timepoint so there are multiple flowers for each tube
# this doesn't include the flower b/c we want a mean
# doesn't use PlantingBlock Either.

# full random model
F_m1 <- lmer(NumShortStamens ~ (1|Population) + (1|Matriline:Population) + (1|Plant:Matriline:Population) + (1|Tube:Plant:Matriline:Population), 
         data = Full)

# mixed model
F_m2 <- lmer(NumShortStamens ~ Population + (1|Matriline:Population) + (1|Plant:Matriline:Population) + (1|Tube:Plant:Matriline:Population), 
         data = Full)

# full fixed model
F_m3 <- lm(NumShortStamens ~ Population + Matriline + Plant + Tube, 
         data = Full)

summary(F_m1)
summary(F_m2)

anova(F_m1)
anova(F_m2)
```

Now just the Sequenced Plants
```{r}

## create a fully nested random model
colnames(Sequenced)
# the important ones are Pop, Line, Rep, Tube to mimic what Jeff used
# Population
# line
# Rep
# Tube
# this doesn't include the flower b/c we want a mean

# full random model
S_m1 <- lmer(Short_Stamens ~ (1|Population) + (1|Line:Population) + (1|Rep:Line:Population) + (1|Tube:Rep:Line:Population), 
         data = Sequenced)

# mixed model
S_m2 <- lmer(Short_Stamens ~ Population + (1|Line:Population) + (1|Rep:Line:Population) + (1|Tube:Rep:Line:Population), 
         data = Sequenced)

# full fixed model
S_m3 <- lm(Short_Stamens ~ Population + Line + Rep + Tube, data = Sequenced)

summary(S_m1)
summary(S_m2)

anova(S_m1)
anova(S_m2)
```

# Try LSMs (estimated marginal means)

Can't use the fully random models here because the package doesn't like that. Deciding to use m2 so most things are random besides the population

```{r}
Full_LSM <- summary(emmeans(F_m2, spec = "Population"))

Seq_LSM <- summary(emmeans(S_m2, spec = "Population"))
```

# Raw Means 

```{r}
Full_RawMeans <- Full %>% group_by(Population) %>%
  summarize(Full_PopFlwrMean = mean(NumShortStamens))

Seq_RawMeans <- Sequenced %>% group_by(Population) %>%
  summarize(Seq_PopFlwrMean = mean(Short_Stamens))


# Create an output file (.ROBJ is good enough I think) with the Population, SSN, and elevation as metadata
Elev_Means <- merge(Full_RawMeans, Seq_RawMeans, by = "Population")
Elev_Means$Elev_m <- PopMetaData$Elev_m[match(Elev_Means$Population, PopMetaData$PopCode)]
save(Elev_Means, file = "C:/Users/Sophie/Michigan State University/Conner, Jeffrey - SophieAnalyses/R_script/StamenLossPipeline/Elev_Means_Feb2022.ROBJ")
```

# Compare values

Here is a comparison of the 4 different means calculated (LSMs and Raw means for both Sequenced and Full datasets)

```{r}
LSMs <- data.frame("Population" = Full_LSM$Population, "Full_LSM" = Full_LSM$emmean, "Pop2" = Seq_LSM$Population, "Seq_LSM" = Seq_LSM$emmean)
Comp <- merge(Elev_Means, LSMs, by = "Population")

# let's see some correlations
cor.test(Comp$Full_PopFlwrMean, Comp$Full_LSM)
cor.test(Comp$Full_PopFlwrMean, Comp$Seq_PopFlwrMean)
cor.test(Comp$Seq_PopFlwrMean, Comp$Seq_LSM)

# so all are highly correlated. Let's make a few plots just to visualize that
plot(Comp$Full_PopFlwrMean ~ Comp$Full_LSM)
lines(x = c(0,1,2), y = c(0,1,2))
plot(Comp$Full_PopFlwrMean ~ Comp$Seq_PopFlwrMean)
lines(x = c(0,1,2), y = c(0,1,2))
plot(Comp$Seq_PopFlwrMean ~ Comp$Seq_LSM)
lines(x = c(0,1,2), y = c(0,1,2))

```
